Reproducibility of COVID-19 pre-prints

### 2022-01-13 outputs/paper/paper.tex
Figure and paper file updates
Address some more comments

\title{Reproducibility of COVID-19 pre-prints\thanks{\textbf{Preprint:} This present manuscript is a version of a preprint by the authors (Collins and Alexander 2021). The text, tables and figures of this manuscript overlap with those of the preprint version by the authors. \textbf{Acknowledgements:} We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Two anonymous reviewers and the editor provided extensive comments that substantially improved this paper and we are grateful to them. \textbf{Funding statement:} Collins thanks CANSSI Ontario and Monica Alexander for financial support. \textbf{Data accessibility statement:} Code and data are available at: \url{https://github.com/anniecollins/reproducibility_markers_in_covid19_preprints}. \textbf{Author contributions:} Collins had the original idea and wrote the first draft of the paper. Collins and Alexander obtained and created the datasets. Collins and Alexander analysed the data. All authors interpreted the data and contributed writing the paper and approved the final version.}}
\author{Annie Collins\footnote{University of Toronto.} \and Rohan Alexander\footnote{University of Toronto, \href{mailto:rohan.alexander@utoronto.ca}{\nolinkurl{rohan.alexander@utoronto.ca}}.}}
\date{11 January 2022}

\begin{document}
\maketitle
\begin{abstract}
To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, medRxiv, and SocArXiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, 79 per cent of those on medRxiv, and 85 per cent of those on SocArXiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields.
\end{abstract}

@@

Scientists use open repositories of papers to disseminate their research more quickly than is possible in traditional journals or conference proceedings, and also to obtain feedback on their work prior to publication. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of scientific communication and a lot of research builds on the pre-prints posted there. Pre-print repositories have been especially important during the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community (Else 2020). The centrality of pre-prints to science means that it is important that the results that are posted are credible. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print.

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories are important. But it is important that scientists impose standards on themselves, and arguably repositories have a role to play here. Following Weissgerber et al. (2021), we examine pre-prints about COVID-19 posted to arXiv, bioRxiv, medRxiv, and SocArXiv from 28 January 2020 through to 30 June 2021. By way of background, each of these four repositories has a different focus: arXiv is general although it has especially high rates of usage from fields like mathematics, physics, and computer science, bioRxiv focuses on biological sciences, medRxiv focuses on health sciences, and SocArXiv focuses on social sciences.

We search for markers of open science as indicators of reproducibility, specifically open data and open code. The definition of reproducibility tends to vary by context and academic field. For the purposes of this paper, we define reproducibility to mean the ability for different researchers to achieve the same results given the same data and computational methods as the original source. This contrasts replicability, which we define as the ability for different researchers to achieve consistent results by conducting the full data collection and analysis process in lieu of reusing original data. These definitions match that of National Academies of Sciences and Medicine (2019) and Cacioppo et al. (2015).

We find that of the papers sampled, approximately 75 per cent of papers from arXiv, 67 per cent of papers from bioRxiv, 79 per cent of papers from medRxiv, and 85 per cent of papers from SocArXiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \ref{fig:flowchart}. Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains approximately the same proportion of papers with these markers.

@@

The remainder of this paper is structured as follows: in Section \ref{methodology} we discuss the process of constructing our dataset through retrieving pre-prints from the arXiv, bioRxiv, medRxiv, and SocArXiv repositories and mining them for open data and open code markers. In Section \ref{results}, we present the results and key findings of this process. Finally, in Section \ref{discussion} we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

@@

Our primary dataset consists of pre-print metadata extracted from the arXiv, bioRxiv, medRxiv, and SocArXiv pre-print repositories via their respective Application Programming Interfaces (APIs). This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date created, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, published DOI (if the pre-print has since been published in a peer-reviewed journal), and download link. The data collection process was conducted separately for COVID-19 and pre-COVID-19 papers.

For COVID-19-related pre-prints, we first created a local copy of each repository containing all metadata for pre-prints posted between 1 January, 2020, and 30 June, 2021. In the case of SocArXiv there are fields that are self-reported by the submitting author about the availability of code and data. We classified individual pre-prints as ``COVID-19-related'' based on whether they contained one or more of the following terms in their title or abstract (case insensitive): ``COVID-19,'' ``COVID 19,'' ``corona virus,'' ``coronavirus,'' ``coronavirus-2,'' ``SARS-CoV-2,'' ``SARSCoV-2,'' and ``2019-nCoV.'' We then randomly sampled pre-prints for further analysis, except for SocArXiv, for which we analyzed all pre-prints that matched the above criteria, because there were only 478.

@@

We checked our sampled pre-prints for open data and code markers using the Open Data Detection in Publications (ODDPub) text mining algorithm (Riedel, Kip, and Bobrov 2020) within the \texttt{oddpub} R package (Riedel 2019). This required downloading each pre-print as a PDF, which we did using \texttt{heapsofpapers} (Alexander and Mahfouz 2021) and then converting the PDFs to text files. Some papers from SocArXiv were only available in Word document format (.docx) and were downloaded as such and then converted to text format. We then conducted the open data and open code detection procedure, which involved searching for keywords and other markers of open data and open code availability. This was conducted using the \texttt{open\_data\_search()} function from the \texttt{oddpub} package. Under original validation process, the ODDPub algorithm performed with a sensitivity of 0.73 and a specificity of 1.00 for open code detection and a sensitivity of 0.73 and a specificity of 0.97 for open data detection compared with manual screening (Riedel, Kip, and Bobrov 2020).

What constitutes open code or open data is complicated and discipline specific. The details of the \texttt{oddpub} approach are available in Riedel, Kip, and Bobrov (2020). The general criteria are that specific mention should be made of where the data and code are located, and that data should be as close to raw as possible. Data and code must also be freely accessible to anyone (no request, application, registration process, or affiliation required). Our work was conducted using the statistical programming language R (R Core Team 2020).

@@

Although SocArXiv's metadata contained author-reported information concerning data and code availability for each pre-print, we elected to use the results of the ODDPub algorithm in our analysis to ensure comparability of results between pre-prints from all repositories.

@@

\caption{Number of pre-prints related to COVID-19 posted to arXiv (N = 4,250), bioRxiv (N = 4,146), medRxiv (N = 13,194), and SocArXiv (N = 478) per month}\label{fig:monthly-papers-total}
\end{figure}

@@

\caption{Number of pre-prints posted per month in 2019 to arXiv (N = 150,018), bioRxiv (N = 31,752), medRxiv (N = 913), and SocArXiv (N = 1,614) per month}\label{fig:2019-results-med-arxiv}
\end{figure}

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase, which may be expected as the repository gained recognition and popularity in the medical research community (Figure \ref{fig:2019-results-med-arxiv}). The number of pre-prints posted monthly to bioRxiv also saw a slight overall increase throughout 2019, while the number of those posted to arXiv and SocArXiv fluctuated throughout the year (Figure \ref{fig:2019-results-med-arxiv}). Due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 21,647 pre-prints were posted to medRxiv between June 2019 and 30 June, 2021, 13,194 of which (approximately 61 per cent) relate to COVID-19.

Of the analyzed pre-prints from 2019, 93 per cent of those posted to arXiv, 63 per cent of those posted to bioRxiv, 75 per cent of those posted to medRxiv, 92 per cent of those posted to SocArXiv showed no indication of open data or open code (Table \ref{tab:2019-published}).

{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/monthly-papers-condition-stack-2019-1}

\caption{Number of sampled pre-prints posted in 2019 from arXiv (n = 1,200), bioRxiv (n = 1,200), medRxiv (n = 913), and SocArXiv (n = 1,200) distinguished by presence of open data or code markers}\label{fig:monthly-papers-condition-stack-2019}
\end{figure}

@@

\caption{\label{tab:2019-published}Counts and proportion of published COVID-19 pre-prints in each repository, 2019}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Repositiory & Published pre-prints & Proportion published\\
\hline
arXiv & 62,045 & 0.41\\
\hline
bioRxiv & 20,445 & 0.64\\
\hline
medRxiv & 555 & 0.61\\
\hline
socArXiv & 605 & 0.37\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:2019-open-publishing}Counts and proportions of open data markers by whether the pre-print was published, 2019}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
Status & Both & Neither & Open code & Open data & Proportion with neither\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{arXiv}}\\
\hspace{1em}Unpublished & 11 & 645 & 44 & 3 & 0.92\\
\hspace{1em}Published & 6 & 467 & 15 & 9 & 0.94\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{bioRxiv}}\\
\hspace{1em}Unpublished & 38 & 264 & 34 & 87 & 0.62\\
\hspace{1em}Published & 78 & 488 & 64 & 147 & 0.63\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{medRxiv}}\\
\hspace{1em}Unpublished & 19 & 277 & 19 & 43 & 0.77\\
\hspace{1em}Published & 38 & 409 & 36 & 72 & 0.74\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{socArXiv}}\\
\hspace{1em}Unpublished & 13 & 659 & 10 & 47 & 0.90\\
\hspace{1em}Published & 10 & 442 & 6 & 13 & 0.94\\
\bottomrule
\end{tabular}
\end{table}

Examining publication rates for pre-pandemic papers, we observe that 41 per cent of pre-prints posted to arXiv, 64 per cent of pre-prints posted to bioRxiv, 61 per cent of pre-prints posted to medRxiv, and 37 per cent of pre-prints posted to SocArXiv during 2019 were eventually peer reviewed and published (Table \ref{tab:2019-published}). When disaggregated by open data and code status, we find that published and unpublished pre-prints contain open data and code markers in similar proportions (Table \ref{tab:2019-open-publishing}).

@@

From the collection of all pre-prints related to COVID-19, we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, 1,000 from bioRxiv, and 478 from SocArXiv. Broadly, we are unable to find markers of either open data or open code for 3,011 pre-prints or approximately 76 per cent of our sample (Table \ref{tab:summarycounts-repositories}). Of the remaining pre-prints, 7 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

When differentiated by repository, we observe that open data and code markers were absent from 75 per cent of the sampled arXiv pre-prints, 67 per cent of the sampled bioRxiv pre-prints, 79 per cent of the sampled medRxiv pre-prints, and 85 per cent of all COVID-19-related SocArXiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \ref{tab:summarycounts-repositories}). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 22 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are similar to McGuinness and Sheppard (2021), who focus on medRxiv and find that 23 per cent describe open data.

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \ref{fig:monthly-papers-condition-stack}). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to any given repository. In our datasets, very few (if any) pre-prints were sampled for the month of January 2020. None of these pre-prints contained open data or open code markers, thus the 0 per cent rate of open data and code for this month across all repositories should be considered an outlier.

@@

On the other hand, the analyzed arXiv and SocArXiv COVID-19-related pre-prints contain a higher proportion of open data and code markers overall than their 2019 counterparts, with a increase of seven per cent in sampled SocArXiv pre-prints and 18 per cent in sampled arXiv papers. For these repositories, the presence of both open data and open code markers was significantly associated with registration during the pandemic, suggesting that pre-prints related to COVID-19 in arXiv and SocArXiv may have more consistently adhered to open science practices than their pre-pandemic counterparts (\(\chi\)\textsuperscript{2} = 93.124, arXiv open data; \(\chi\)\textsuperscript{2} = 106.88, arXiv open code; \(\chi\)\textsuperscript{2} = 12.303, SocArXiv open data; \(\chi\)\textsuperscript{2} = 8.1598, SocArXiv open code; p \textless{} 0.005).

@@

{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/monthly-papers-condition-stack-1}

@@

\caption{Number of sampled pre-prints related to COVID-19 from arXiv (n = 1,000), bioRxiv (n = 1,000), medRxiv (n = 1,500), and SocArXiv (n = 478) distinguished by presence of open data or code markers}\label{fig:monthly-papers-condition-stack}
\end{figure}

@@

The proportion of pre-prints that have been published varies by repository (Table \ref{tab:published-sample-summary}). Notably, of all COVID-19-related pre-prints in our dataset, approximately 30 per cent of those posted to bioRxiv and nearly one third of those posted to medRxiv were published. This is high in comparison to the proportions from arXiv and SocArXiv, and although this might suggest that COVID-19-related pre-prints in biomedical fields have received greater attention overall than pre-prints from other fields, our results in Section \ref{pre-pandemic-pre-prints} suggest that this pattern pre-dates the pandemic.

@@

\caption{\label{tab:published-sample-summary}Counts and proportion of published COVID-19 pre-prints in each repository}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Repository & Published pre-prints & Proportion published\\
\hline
arXiv & 717 & 0.17\\
\hline
bioRxiv & 1,230 & 0.30\\
\hline
medRxiv & 4,230 & 0.32\\
\hline
SocArXiv & 71 & 0.15\\
\hline
\end{tabular}
\end{table}

@@

There is limited literature examining the relationship between data and code availability in manuscripts between the pre-print and publication stages. McGuinness and Sheppard (2021) examine differences in data availability statements between medRxiv pre-prints and their published counterparts. They find that data availability was maintained for the vast majority of their sample, varying by journal data sharing policy with greater improvements in openness among manuscripts published in journals mandating data sharing. While limited to medRxiv, McGuinness and Sheppard's results mimic our own work and provide initial evidence to suggest that data availability is generally maintained or improved between the pre-print and publication stages.

@@

\caption{\label{tab:open-published-summary}Counts and proportions of open data markers by whether the pre-print was published}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
Status & Both & Neither & Open code & Open data & Proportion with neither\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{arXiv}}\\
\hspace{1em}Unpublished & 79 & 622 & 100 & 27 & 0.75\\
\hspace{1em}Published & 20 & 130 & 15 & 7 & 0.76\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{bioRxiv}}\\
\hspace{1em}Unpublished & 55 & 471 & 29 & 145 & 0.67\\
\hspace{1em}Published & 20 & 203 & 15 & 62 & 0.68\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{medRxiv}}\\
\hspace{1em}Unpublished & 10 & 112 & 8 & 8 & 0.81\\
\hspace{1em}Published & 113 & 1,068 & 81 & 100 & 0.78\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{SocArXiv}}\\
\hspace{1em}Unpublished & 14 & 345 & 14 & 34 & 0.85\\
\hspace{1em}Published & 3 & 60 & 0 & 8 & 0.85\\
\bottomrule
\end{tabular}
\end{table}

@@

Open data is generally accepted to be beneficial to the scientific process and to a paper's reproducibility potential, hence it is concerning that around 75 per cent of pre-prints in our sample contained no open data markers. This concern is slightly mitigated by recognition of challenges in working with biomedical data compared with data in other fields, notably privacy and ethics concerns when working with personal data (Floca 2014). The COVID-19 pandemic has seen open science initiatives, as evidenced by the creation of open data repositories such as the dashboard maintained by the Center for Systems Science and Engineering at Johns Hopkins University (Dong, Du, and Gardner 2020) or the large number of publishers who have removed paywalls from published COVID-19 research (Gill 2020). While the intention at the start of the pandemic was that there would be `clear statements regarding the availability of underlying data' (Wellcome Trust 2020) some retractions of work have been based on `unreliable or nonexistent data' (Teixeira da Silva, Bornemann-Cimenti, and Tsigaris 2021). SocArXiv asks submitters to make explicit the degree of openness in the data, and this is something that could be adopted by other pre-print repositories.

@@

\hypertarget{tables}{%
\section{Tables}\label{tables}}

>>>>

\title{Reproducibility of COVID-19 pre-prints\thanks{\textbf{Preprint:} This present manuscript is a version of a preprint by the authors (Collins and Alexander 2021). The text, tables and figures of this manuscript overlap with those of the preprint version by the authors. \textbf{Acknowledgements:} We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Two anonymous reviewers and the editor provided extensive comments that substantially improved this paper and we are grateful to them. \textbf{Funding statement:} Collins thanks CANSSI Ontario and Monica Alexander for financial support. \textbf{Data accessibility statement:} Code and data are available at: \url{https://github.com/anniecollins/reproducibility_markers_in_covid19_preprints}. \textbf{Author contributions:} Collins had the original idea and wrote the first draft of the paper. Collins and Alexander obtained and created the datasets. Collins and Alexander analysed the data. All authors interpreted the data and contributed writing the paper and approved the final version.}}
\author{Annie Collins\footnote{University of Toronto.} \and Rohan Alexander\footnote{University of Toronto, \href{mailto:rohan.alexander@utoronto.ca}{\nolinkurl{rohan.alexander@utoronto.ca}}.}}
\date{13 January 2022}

\begin{document}
\maketitle
\begin{abstract}
To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, and medRxiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, and 79 per cent of those on medRxiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields.
\end{abstract}

@@

Scientists use open repositories of papers to disseminate their research more quickly than is possible in traditional journals or conference proceedings, and also to obtain feedback on their work prior to publication. These repositories, such as arXiv, bioRxiv, and medRxiv, are a critical component of scientific communication and a lot of research builds on the pre-prints posted there. Pre-print repositories have been especially important during the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community (Else 2020). The centrality of pre-prints to science means that it is important that the results that are posted are credible. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print.

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories are important. But it is important that scientists impose standards on themselves, and arguably repositories have a role to play here. Following Weissgerber et al. (2021), we examine pre-prints about COVID-19 posted to arXiv, bioRxiv, and medRxiv from 28 January 2020 through to 30 June 2021. By way of background, each of these three repositories has a different focus: arXiv is general although it has especially high rates of usage from fields like mathematics, physics, and computer science, bioRxiv focuses on biological sciences, and medRxiv focuses on health sciences.

We search for markers of open science as indicators of reproducibility, specifically open data and open code. The definition of reproducibility tends to vary by context and academic field (Barba 2018). For the purposes of this paper, we define reproducibility to mean the ability for different researchers to achieve the same results given the same data and computational methods as the original source. This contrasts replicability, which we define as the ability for different researchers to achieve consistent results by conducting the full data collection and analysis process in lieu of reusing original data. These definitions match that of National Academies of Sciences and Medicine (2019) and Cacioppo et al. (2015).

We find that of the papers sampled, approximately 75 per cent of papers from arXiv, 67 per cent of papers from bioRxiv, and 79 per cent of papers from medRxiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \ref{fig:flowchart}. Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains approximately the same proportion of papers with these markers.

@@

The remainder of this paper is structured as follows: in Section \ref{methodology} we discuss the process of constructing our dataset through retrieving pre-prints from the arXiv, bioRxiv, and medRxiv repositories and mining them for open data and open code markers. In Section \ref{results}, we present the results and key findings of this process. Finally, in Section \ref{discussion} we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

@@

Our primary dataset consists of pre-print metadata extracted from the arXiv, bioRxiv, and medRxiv pre-print repositories via their respective Application Programming Interfaces (APIs). This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date created, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, published DOI (if the pre-print has since been published in a peer-reviewed journal), and download link. The data collection process was conducted separately for COVID-19 and pre-COVID-19 papers.

For COVID-19-related pre-prints, we first created a local copy of each repository containing all metadata for pre-prints posted between 1 January, 2020, and 30 June, 2021.
We classified individual pre-prints as ``COVID-19-related'' based on whether they contained one or more of the following terms in their title or abstract (case insensitive): ``COVID-19,'' ``COVID 19,'' ``corona virus,'' ``coronavirus,'' ``coronavirus-2,'' ``SARS-CoV-2,'' ``SARSCoV-2,'' and ``2019-nCoV.'' We then randomly sampled pre-prints for further analysis.

@@

We checked our sampled pre-prints for open data and code markers using the Open Data Detection in Publications (ODDPub) text mining algorithm (Riedel, Kip, and Bobrov 2020) within the \texttt{oddpub} R package (Riedel 2019) (RRID:SCR\_018385) . This required downloading each pre-print as a PDF and then converting the PDFs to text files.
We then conducted the open data and open code detection procedure, which involved searching for keywords and other markers of open data and open code availability. This was conducted using the \texttt{open\_data\_search()} function from the \texttt{oddpub} package. Under original validation process, the ODDPub algorithm performed with a sensitivity of 0.73 and a specificity of 1.00 for open code detection and a sensitivity of 0.73 and a specificity of 0.97 for open data detection compared with manual screening (Riedel, Kip, and Bobrov 2020).

What constitutes open code or open data is complicated and discipline specific. The details of the \texttt{oddpub} approach are available in Riedel, Kip, and Bobrov (2020). The general criteria are that specific mention should be made of where the data and code are located, and that data should be as close to raw as possible. Data and code must also be freely accessible to anyone (no request, application, registration process, or affiliation required). Our work was conducted using the statistical programming language R (R Core Team 2020) (RRID:SCR\_001905).

@@

@@

\caption{Number of pre-prints related to COVID-19 posted to arXiv (N = 4,250), bioRxiv (N = 4,146), and medRxiv (N = 13,194) per month}\label{fig:monthly-papers-total}
\end{figure}

@@

\caption{Number of pre-prints posted per month in 2019 to arXiv (N = 150,018), bioRxiv (N = 31,752), medRxiv (N = 913)}\label{fig:2019-results-med-arxiv}
\end{figure}

Between June and December 2019, the number of pre-prints posted to medRxiv monthly saw an overall increase, which may be expected as the repository gained recognition and popularity in the medical research community (Figure \ref{fig:2019-results-med-arxiv}). The number of pre-prints posted monthly to bioRxiv also saw a slight overall increase throughout 2019, while the number of those posted to arXiv fluctuated throughout the year (Figure \ref{fig:2019-results-med-arxiv}). Due to its relative immaturity at the beginning of the COVID-19 pandemic, a significant portion of medRxiv's overall usage has been dedicated to COVID-19-related research. In total, 21,647 pre-prints were posted to medRxiv between June 2019 and 30 June, 2021, 13,194 of which (approximately 61 per cent) relate to COVID-19.

Of the analyzed pre-prints from 2019, 93 per cent of those posted to arXiv, 63 per cent of those posted to bioRxiv, and 75 per cent of those posted to medRxiv showed no indication of open data or open code (Table \ref{tab:2019-published}).

@@

\caption{\label{tab:2019-published}Counts and proportion of published COVID-19 pre-prints in each repository, 2019}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Repositiory & Published pre-prints & Proportion published\\
\hline
arXiv & 62,045 & 0.41\\
\hline
bioRxiv & 20,445 & 0.64\\
\hline
medRxiv & 555 & 0.61\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:2019-open-publishing}Counts and proportions of open data markers by whether the pre-print was published, 2019}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
Status & Both & Neither & Open code & Open data & Proportion with neither\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{arXiv}}\\
\hspace{1em}Unpublished & 11 & 645 & 44 & 3 & 0.92\\
\hspace{1em}Published & 6 & 467 & 15 & 9 & 0.94\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{bioRxiv}}\\
\hspace{1em}Unpublished & 38 & 264 & 34 & 87 & 0.62\\
\hspace{1em}Published & 78 & 488 & 64 & 147 & 0.63\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{medRxiv}}\\
\hspace{1em}Unpublished & 19 & 277 & 19 & 43 & 0.77\\
\hspace{1em}Published & 38 & 409 & 36 & 72 & 0.74\\
\bottomrule
\end{tabular}
\end{table}

Examining publication rates for pre-pandemic papers, we observe that 41 per cent of pre-prints posted to arXiv, 64 per cent of pre-prints posted to bioRxiv, and 61 per cent of pre-prints posted to medRxiv during 2019 were eventually peer reviewed and published (Table \ref{tab:2019-published}). When disaggregated by open data and code status, we find that published and unpublished pre-prints contain open data and code markers in similar proportions (Table \ref{tab:2019-open-publishing}).

@@

From the collection of all pre-prints related to COVID-19, we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, and 1,000 from bioRxiv. Broadly, we are unable to find markers of either open data or open code for 3,011 pre-prints or approximately 76 per cent of our sample (Table \ref{tab:summarycounts-repositories}). Of the remaining pre-prints, 7 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

When differentiated by repository, we observe that open data and code markers were absent from 75 per cent of the sampled arXiv pre-prints, 67 per cent of the sampled bioRxiv pre-prints, 79 per cent of the sampled medRxiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \ref{tab:summarycounts-repositories}). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 22 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are similar to McGuinness and Sheppard (2021), who focus on medRxiv and find that 23 per cent describe open data.

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \ref{fig:arxiv-and-monthly-papers-condition-stack}). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to any given repository. In our datasets, very few (if any) pre-prints were sampled for the month of January 2020. None of these pre-prints contained open data or open code markers, thus the 0 per cent rate of open data and code for this month across all repositories should be considered an outlier.

It is also important to note that pre-prints posted during the early months of the pandemic were likely using, and reusing, publicly available data sources due to an inability to collect original data within a short timeframe. Additionally, \texttt{oddpub} does not consider `{[}t{]}he reuse of data/code previously published by other researchers' (Riedel, Kip, and Bobrov 2020). A different definition of open data could enable pre-prints that reuse publicly available data to have made theirdata available for reproducibility purposes.

@@

On the other hand, the analyzed arXiv COVID-19-related pre-prints contain a higher proportion of open data and code markers overall than their 2019 counterparts, with a increase of 18 per cent in sampled arXiv papers. For these repositories, the presence of both open data and open code markers was significantly associated with registration during the pandemic, suggesting that pre-prints related to COVID-19 in arXiv may have more consistently adhered to open science practices than their pre-pandemic counterparts (\(\chi\)\textsuperscript{2} = 93.124, arXiv open data; \(\chi\)\textsuperscript{2} = 106.88, arXiv open code; \(\chi\)\textsuperscript{2} = 12.303).

@@

{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/arxiv-and-monthly-papers-condition-stack-1}

@@

\caption{Number of sampled pre-prints posted in 2019, compared with those related to COVID-19: from arXiv (n in 2019 = 1,000; n related to COVID = 1,200), bioRxiv (n in 2019 = 1,000; n related to COVID n = 1,200), and medRxiv (n in 2019 = 1,500; n related to COVID n = 913) distinguished by presence of open data or code markers}\label{fig:arxiv-and-monthly-papers-condition-stack}
\end{figure}

@@

The proportion of pre-prints that have been published varies by repository (Table \ref{tab:published-sample-summary}). Notably, of all COVID-19-related pre-prints in our dataset, approximately 30 per cent of those posted to bioRxiv and nearly one third of those posted to medRxiv were published. This is high in comparison to the proportions from arXiv, and although this might suggest that COVID-19-related pre-prints in biomedical fields have received greater attention overall than pre-prints from other fields, our results in Section \ref{pre-pandemic-pre-prints} suggest that this pattern pre-dates the pandemic.

@@

\caption{\label{tab:published-sample-summary}Counts and proportion of published COVID-19 pre-prints in each repository}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Repository & Published pre-prints & Proportion published\\
\hline
arXiv & 717 & 0.17\\
\hline
bioRxiv & 1,230 & 0.30\\
\hline
medRxiv & 4,230 & 0.32\\
\hline
\end{tabular}
\end{table}

@@

There is limited literature examining the relationship between data and code availability in manuscripts between the pre-print and publication stages. McGuinness and Sheppard (2021) examine differences in data availability statements between medRxiv pre-prints and their published counterparts. They find that data availability was maintained for the vast majority of their sample, varying by journal data sharing policy with greater improvements in openness among manuscripts published in journals mandating data sharing. While limited to medRxiv, the results of McGuinness and Sheppard (2021) align with our own work and provide initial evidence to suggest that data availability is generally maintained or improved between the pre-print and publication stages.

@@

\caption{\label{tab:open-published-summary}Counts and proportions of open data markers by whether the pre-print was published}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
Status & Both & Neither & Open code & Open data & Proportion with neither\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{arXiv}}\\
\hspace{1em}Unpublished & 79 & 622 & 100 & 27 & 0.75\\
\hspace{1em}Published & 20 & 130 & 15 & 7 & 0.76\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{bioRxiv}}\\
\hspace{1em}Unpublished & 55 & 471 & 29 & 145 & 0.67\\
\hspace{1em}Published & 20 & 203 & 15 & 62 & 0.68\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{medRxiv}}\\
\hspace{1em}Unpublished & 10 & 112 & 8 & 8 & 0.81\\
\hspace{1em}Published & 113 & 1,068 & 81 & 100 & 0.78\\
\bottomrule
\end{tabular}
\end{table}

@@

Open data is generally accepted to be beneficial to the scientific process and to a paper's reproducibility potential, hence it is concerning that around 75 per cent of pre-prints in our sample contained no open data markers. This concern is slightly mitigated by recognition of challenges in working with biomedical data compared with data in other fields, notably privacy and ethics concerns when working with personal data (Floca 2014). The COVID-19 pandemic has seen open science initiatives, as evidenced by the creation of open data repositories such as the dashboard maintained by the Center for Systems Science and Engineering at Johns Hopkins University (Dong, Du, and Gardner 2020) or the large number of publishers who removed paywalls from published COVID-19 research (Gill 2020). While the intention at the start of the pandemic was that there would be `clear statements regarding the availability of underlying data' (Wellcome Trust 2020) some retractions of work have been based on `unreliable or nonexistent data' (Teixeira da Silva, Bornemann-Cimenti, and Tsigaris 2021).

@@

\hypertarget{tables-with-specific-values-displayed-in-the-graphs}{%
\section{Tables with specific values displayed in the graphs}\label{tables-with-specific-values-displayed-in-the-graphs}}

\begin{table}[!h]
